Design of Adaptive Constellations and Error Protection Coding for Wireless Network Coding in 5-node Butterfly Networks

### 2017-05-25 Const_design.tex
Redrawn Figures in 3-step comparison and added some comments.

For the sake of simplicity of the theoretical performance analysis,
we assume that source$\rightarrow$relay channels in WBN are AWGN\footnote{Note that this also includes the system with perfect source phase
prerotation \cite{Koike-Popovski-Tarokh_2009a,Zhang-Liew_2010,Anxin-Yuan-Kayama_2008},
which can be obtained by a generalization of the scaling coefficients
as $L'_{n}=L_{n}e^{-\jj\measuredangle h_{SR}}$, where $h_{SR}$ is
the corresponding source$\rightarrow$relay channel.} (i.e. $h_{iR}=1$ for $i\in\{A,B\}$). We evaluate the throughput
analytically as $(N_{b}+N_{s})\cdot(1-P_{e}^{\mathrm{FER}})$, where
$P_{e}^{\mathrm{FER}}$ denotes the frame error rate. Due to the system
symmetry we evaluate the throughput of $S_{A}\to D_{A}$ only. We
set the frame length to $M=768$ symbols (to match the HW evaluation)
and the relay output mapper $\mathcal{A}^{R}$ will be defined as
a conventional ($2^{N_{b}+2N_{s}}$$)$-QAM constellation. All performance
metrics in the symmetric WBN are evaluated as a function of the actual
SNRs ($\gamma_{{\rm {MAC}}},\,\gamma_{{\rm {HSI}}},\,\gamma_{{\rm {BC}}}$).
\begin{figure*}
\centering{}
\includegraphics[width=0.8\textwidth]{fig/d2_1}
\caption{Source constellation design example for $N_{b}=2,\,N_{s}=1$. Resulting
constellations are depicted as blue circles ($S_{A}$ output constellation),
red circles ($S_{B}$ output constellation) and squares (received
superimposed constellation at $R$). Hierarchical function is $f\left(\mathbf{d}_{A}^{b},\,\mathbf{d}_{B}^{b}\right)=\mathbf{d}_{A}^{b}\oplus\mathbf{d}_{B}^{b}$.\label{fig:CTUpp_ConstellationDesignExample} }
\end{figure*}
\begin{figure*}
\begin{centering}
\includegraphics[width=0.95\textwidth]{fig/const_all_2}
\par\end{centering}

@@

The desired adaptive system parameters can be obtained in the following
two-step procedure. First of all, we analyze the cut-set bound rate
regions $\mathcal{R}^{(N_{b},N_{s})}$ (as discussed in Section \ref{sub:CTUpp_Transmission-Rate-Region})
to identify the maximal encoded system throughput $T_{C}^{(N_{b},N_{s})}$
for a fixed source constellation $(N_{b},N_{s})$:
\begin{align}
T_{C}^{(N_{b},N_{s})}(\gamma_{{\rm {MAC}}},\,\gamma_{{\rm {BC}}},\,\gamma_{{\rm {HSI}}}) & =\max_{r_{b},r_{s}\in\mathcal{R}^{(N_{b},N_{s})}}(r_{b}+r_{s}),\\
 & =r_{b}^{\mathrm{max}}(N_{b},N_{s})+r_{s}^{\mathrm{max}}(N_{b},N_{s}).\label{eq:CTUpp_opt_rates}
\end{align}
 Then, an exhaustive search over the constellation parameters $N_{b},\,N_{s}$
can be performed to find the maximum encoded system throughput $T_{C}^{\mathrm{max}}$
for the given SNR conditions ($\gamma_{{\rm {MAC}}},\,\gamma_{{\rm {BC}}},\,\gamma_{{\rm {HSI}}}$):
\begin{equation}
T_{C}^{\mathrm{max}}(\gamma_{{\rm {MAC}}},\,\gamma_{{\rm {BC}}},\,\gamma_{{\rm {HSI}}})=T_{C}^{(N_{b}^{\mathrm{II}},\,N_{s}^{\mathrm{II}})}=\max_{(N_{b},N_{s})}T_{C}^{(N_{b},N_{s})},\label{eq:CTUpp_Throughput_encoded}
\end{equation}
where $(N_{b}^{\mathrm{II}},\,N_{s}^{\mathrm{II}})$ are the \emph{optimal
source constellation parameters}. The corresponding \emph{optimal
source transmission rates} (\ref{eq:CTUpp_opt_rates}) are given simply
by \emph{$r_{b}^{\mathrm{II}}=r_{b}^{\mathrm{max}}(N_{b}^{\mathrm{II}},\,N_{s}^{\mathrm{II}})$,
$r_{s}^{\mathrm{II}}=r_{s}^{\mathrm{max}}(N_{b}^{\mathrm{II}},\,N_{s}^{\mathrm{II}})$.}
\begin{figure*}
\centering{}
\includegraphics[width=0.45\textwidth]{fig/Adaptive_Maps_overall_BC_7\lyxdot 5}~~~
\includegraphics[width=0.45\textwidth]{fig/Adaptive_Maps_overall_BC_15\lyxdot 0}
\caption{Maximal encoded system throughput $T_{C}^{\mathrm{max}}$ for $\gamma_{{\rm {BC}}}=7.5\,\mathrm{dB}$
and $\gamma_{{\rm {BC}}}=15\,\mathrm{dB}$.\label{fig:CTUpp_Coded_SNR_map} }
\end{figure*}
\begin{figure*}
\centering{}
\includegraphics[width=0.45\textwidth]{fig/XOR_map_Througput_coding_Gain7\lyxdot 5}~~~
\includegraphics[width=0.45\textwidth]{fig/XOR_map_Througput_coding_Gain15\lyxdot 0}
\caption{Throughput enhancement $\Delta T_{C}=T_{C}^{\mathrm{max}}-T_{\mathrm{LB}}$
{[}bits/channel use{]} of the coded over uncoded WBN system for $\gamma_{{\rm {BC}}}=7.5\,\mathrm{dB}$
and $\gamma_{{\rm {BC}}}=15\,\mathrm{dB}$. Note that $0.75\leq\Delta T_{C}\leq2$
in the analyzed range of channel SNRs.\label{fig:CTUpp_Coded_ThroughputGain}}
\end{figure*}

@@

\subsection{Robustness Analysis}

@@

\subsection{Three-Step Reference Scenario}
For a more relevant comparison of the proposed method, we consider two conventional \emph{3-step}
communication scenarios.

The first scenario is equivalent to a conventional routing, where the sources subsequently transmit
their data packet towards the relay. Relay then sends a complete source information (no NC coding is
applied) and thus the HSI channels are completely ignored.

In the second scenario both sources again transmit their information one by one, but the relay is
allowed to combine the received (fully decoded) data packets into the network coded packet
$f(d_{A},d_{B})=d_{A}\oplus d_{B}$ and HSI channels are necessary to retrieve the desired
information at both sources.

@@

\begin{figure}
\begin{centering}
\includegraphics[width=0.75\columnwidth]{fig/scheme_a}
\par\end{centering}
\centering{}\caption{Performance comparison .\label{fig:CTUpp_comparison-schemaa}}
\end{figure}

\begin{figure}
\begin{centering}
\includegraphics[width=0.75\columnwidth]{fig/scheme_b}
\par\end{centering}
\centering{}\caption{Performance comparison .\label{fig:CTUpp_comparison-schemab}}
\end{figure}

@@

In future work, we would like to focus on a full adaptive HW implementation
of the proposed system, where the impact of several other parameters
(e.g. channel fading and dynamics, asymmetry of channel gains etc.)
can significantly affect the system performance and possibly even
require some modifications of the adaptive system design.
\begin{figure*}
\centering{\includegraphics[width=1\textwidth]{fig/Robust_20}
\includegraphics[width=1\textwidth]{fig/Robust_21}}
\caption{Robustness analysis for two fixed $(N_{b},N_{s},r_{b},r_{s})$ scenarios.
\label{fig:CTUpp_Robustenss_analysis}}
\end{figure*}

>>>>

For the sake of simplicity of the theoretical performance analysis,
we assume that source$\rightarrow$relay channels in WBN are AWGN\footnote{Note that this also includes the system with perfect source phase
prerotation \cite{Koike-Popovski-Tarokh_2009a,Zhang-Liew_2010,Anxin-Yuan-Kayama_2008},
which can be obtained by a generalization of the scaling coefficients
as $L'_{n}=L_{n}e^{-\jj\measuredangle h_{SR}}$, where $h_{SR}$ is
the corresponding source$\rightarrow$relay channel.} (i.e. $h_{iR}=1$ for $i\in\{A,B\}$). We evaluate the throughput
analytically as $(N_{b}+N_{s})\cdot(1-P_{e}^{\mathrm{FER}})$, where
$P_{e}^{\mathrm{FER}}$ denotes the frame error rate. Due to the system
symmetry we evaluate the throughput of $S_{A}\to D_{A}$ only. We
set the frame length to $M=768$ symbols (to match the HW evaluation)
and the relay output mapper $\mathcal{A}^{R}$ will be defined as
a conventional $(2^{N_{b}+2N_{s}})$-QAM constellation. All performance
metrics in the symmetric WBN are evaluated as a function of the actual
SNRs ($\gamma_{{\rm {MAC}}},\,\gamma_{{\rm {HSI}}},\,\gamma_{{\rm {BC}}}$).
\begin{figure*}
\centering{}
\includegraphics[width=0.8\textwidth]{fig/d2_1}
\caption{Source constellation design example for $N_{b}=2,\,N_{s}=1$. Resulting
constellations are depicted as blue circles ($S_{A}$ output constellation),
red circles ($S_{B}$ output constellation) and squares (received
superimposed constellation at $R$). Hierarchical function is $f\left(\mathbf{d}_{A}^{b},\,\mathbf{d}_{B}^{b}\right)=\mathbf{d}_{A}^{b}\oplus\mathbf{d}_{B}^{b}$.\label{fig:CTUpp_ConstellationDesignExample} }
\end{figure*}
\begin{figure*}
\begin{centering}
\includegraphics[width=0.95\textwidth]{fig/const_all_2}
\par\end{centering}

@@

The desired adaptive system parameters can be obtained in the following
two-step procedure. First of all, we analyze the cut-set bound rate
regions $\mathcal{R}^{(N_{b},N_{s})}$ (as discussed in Section \ref{sub:CTUpp_Transmission-Rate-Region})
to identify the maximal encoded system throughput $T_{C}^{(N_{b},N_{s})}$
for a fixed source constellation $(N_{b},N_{s})$:
\begin{align}
T_{C}^{(N_{b},N_{s})}(\gamma_{{\rm {MAC}}},\,\gamma_{{\rm {BC}}},\,\gamma_{{\rm {HSI}}}) & =\max_{r_{b},r_{s}\in\mathcal{R}^{(N_{b},N_{s})}}(r_{b}+r_{s}),\\
 & =r_{b}^{\mathrm{max}}(N_{b},N_{s})+r_{s}^{\mathrm{max}}(N_{b},N_{s}).\label{eq:CTUpp_opt_rates}
\end{align}
 Then, an exhaustive search over the constellation parameters $N_{b},\,N_{s}$
can be performed to find the maximum encoded system throughput $T_{C}^{\mathrm{max}}$
for the given SNR conditions ($\gamma_{{\rm {MAC}}},\,\gamma_{{\rm {BC}}},\,\gamma_{{\rm {HSI}}}$):
\begin{equation}
T_{C}^{\mathrm{max}}(\gamma_{{\rm {MAC}}},\,\gamma_{{\rm {BC}}},\,\gamma_{{\rm {HSI}}})=T_{C}^{(N_{b}^{\mathrm{II}},\,N_{s}^{\mathrm{II}})}=\max_{(N_{b},N_{s})}T_{C}^{(N_{b},N_{s})},\label{eq:CTUpp_Throughput_encoded}
\end{equation}
where $(N_{b}^{\mathrm{II}},\,N_{s}^{\mathrm{II}})$ are the \emph{optimal
source constellation parameters}. The corresponding \emph{optimal
source transmission rates} (\ref{eq:CTUpp_opt_rates}) are given simply
by \emph{$r_{b}^{\mathrm{II}}=r_{b}^{\mathrm{max}}(N_{b}^{\mathrm{II}},\,N_{s}^{\mathrm{II}})$,
$r_{s}^{\mathrm{II}}=r_{s}^{\mathrm{max}}(N_{b}^{\mathrm{II}},\,N_{s}^{\mathrm{II}})$.}
\begin{figure*}
\centering{}
\includegraphics[width=0.45\textwidth]{fig/Adaptive_Maps_overall_BC_7\lyxdot 5}~~~
\includegraphics[width=0.45\textwidth]{fig/Adaptive_Maps_overall_BC_15\lyxdot 0}
\caption{Maximal encoded system throughput $T_{C}^{\mathrm{max}}$ for $\gamma_{{\rm {BC}}}=7.5\,\mathrm{dB}$
and $\gamma_{{\rm {BC}}}=15\,\mathrm{dB}$.\label{fig:CTUpp_Coded_SNR_map} }
\end{figure*}
\begin{figure*}
\centering{}
\includegraphics[width=0.45\textwidth]{fig/XOR_map_Througput_coding_Gain7\lyxdot 5}~~~
\includegraphics[width=0.45\textwidth]{fig/XOR_map_Througput_coding_Gain15\lyxdot 0}
\caption{Throughput enhancement $\Delta T_{C}=T_{C}^{\mathrm{max}}-T_{\mathrm{LB}}$
[bits/channel use] of the coded over uncoded WBN system for $\gamma_{{\rm {BC}}}=7.5\,\mathrm{dB}$
and $\gamma_{{\rm {BC}}}=15\,\mathrm{dB}$. Note that $0.75\leq\Delta T_{C}\leq2$
in the analyzed range of channel SNRs.\label{fig:CTUpp_Coded_ThroughputGain}}
\end{figure*}

@@

\subsection{Robustness Analysis}\label{sub:Robustness_Analysis}

@@

\begin{figure*}
\centering{\includegraphics[width=1\textwidth]{fig/Robust_20}
\includegraphics[width=1\textwidth]{fig/Robust_21}}
\caption{Robustness analysis for two fixed $(N_{b},N_{s},r_{b},r_{s})$ scenarios.
\label{fig:CTUpp_Robustenss_analysis}}
\end{figure*}

\subsection{Three-Step Reference Scenario}
For a more advanced comparison of the proposed method, we consider two conventional \emph{3-step}
communication scenarios.

The first scenario is equivalent to a conventional routing, where the sources subsequently transmit
their data packet towards the relay. Relay then sends a complete source information (no NC coding is
applied) and thus the HSI channels are completely ignored (see Figure \ref{fig:CTUpp_ref3stepRouting}).

In the second scenario both sources again transmit their information one by one, but the relay is
allowed to combine the received (fully decoded) data packets into the network coded packet
$f(d_{A},d_{B})=d_{A}\oplus d_{B}$ and HSI channels are necessary to retrieve the desired
information at both sources (see Figure \ref{fig:CTUpp_ref3stepNC}).

@@

We compared throughput of the proposed adaptive constellation design from Section
\ref{sub:Maximal-Throughput-of} on the USRP devices with optimized tuple ($r_b, r_s, N_b, N_s$) for
each measured point (HW-pre-rotation on) with the reference scenarios in Figures \ref{fig:CTUpp_comparison-schemaa} and
\ref{fig:CTUpp_comparison-schemab}. We also considered the proposed method with
pre-rotation turned off as in Section \ref{sub:Robustness_Analysis} for comparison. Finally, we
added two theoretical limits for the 3-step scenarios, where ideal Gaussian alphabets are assumed.

\begin{figure}
\begin{centering}
\includegraphics[width=0.75\columnwidth]{fig/scheme_a}
\par\end{centering}
\centering{}\caption{Performance comparison for the setup
$\gamma_{\mathrm{BC}}\approx\gamma_{\mathrm{MAC}}$ and
$\gamma_{\mathrm{HSI}}\approx\gamma_{\mathrm{MAC}}$+3dB
.\label{fig:CTUpp_comparison-schemaa}}
\end{figure}

\begin{figure}
\begin{centering}
\includegraphics[width=0.75\columnwidth]{fig/scheme_b}
\par\end{centering}
\centering{}\caption{Performance comparison for the setup
$\gamma_{\mathrm{BC}}\approx\gamma_{\mathrm{MAC}}$+5dB and
$\gamma_{\mathrm{HSI}}\approx\gamma_{\mathrm{MAC}}$-1dB
.\label{fig:CTUpp_comparison-schemab}}
\end{figure}

One can observe that

@@

In future work, we would like to focus on a full adaptive HW implementation
of the proposed system, where the impact of several other parameters
(e.g. channel fading and dynamics, asymmetry of channel gains etc.)
can significantly affect the system performance and possibly even
require some modifications of the adaptive system design.
